The GWR Smoke's effect on Oregon 2013 Medication of Beta-2 agonists

### 2017-08-04 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

@@

smoke_map1 <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
  # fill with number of smoke days
  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
  scale_fill_gradient("Smoke-Impacted Days",
                      low = "#3964C3", high = "#dc2430") +
  # add county path on top
  geom_path() +
  facet_wrap(~smk_cut_method) +
   theme(strip.background = element_blank(),
       strip.text.x = element_blank()) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(colour=NA, fill=NA)) +
  geom_text(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2)
smoke_map

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- inhaler_ndc_df %>%
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # only look at outcomes
  filter(outcome == 1) %>%
  # add another row that makes sure there is a person <15 in the dataframe
  # tricking xtabs to produce a 0 cell for the outcome for age <15
  add_row(outcome = 0, age_ind = 0)

@@

### Outcomes

```{r category, echo = F, message = F, warning = F, results='asis', eval=FALSE}
read_path <- paste0("../data_new/medication/oregon_beta_ndc_index_full.csv")
inhaler_ndc <- read_csv(read_path)

@@

@@

### Find the smallest of AIC with df from 3 to 6

```{r, message=FALSE, echo=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

@@

>>>>

centroids[4,] <- centroids[4,] + c(0, 0.02) # Clatsop
centroids[5,] <- centroids[5,] + c(0.1, 0) # Columbia
centroids[29,] <- centroids[29,] + c(-0.1, -0.03) # Tillamook
centroids[34,] <- centroids[34,] + c(-0.12, -0.02) # Washington
centroids[26,] <- centroids[26,] + c(0, 0.03) # Multnomah
centroids[11,] <- centroids[11,] + c(-0.05, -0.12) # Gilliam
centroids[24,] <- centroids[24,] + c(-0.2, 0) # Marion
centroids[33,] <- centroids[33,] + c(0, -0.1) # Wasco
centroids[28,] <- centroids[28,] + c(0, 0.1) # Sherman
centroids[36,] <- centroids[36,] + c(-0.05, -0.08) # Yamhill
centroids[8,] <- centroids[8,] + c(0.08, -0.2) # Yamhill
centroids[17,] <- centroids[17,] + c(-0.05, 0.08) # Yamhill
centroids[29,] <- centroids[29,] + c(-0.01, -0.1) # Sherman

@@

smoke_map1 <- ggplot(or_smk_df, aes(x=long, y=lat, group=group)) +
  # fill with number of smoke days
  geom_polygon(aes(fill = smoke_days), alpha = 0.7) +
  scale_fill_gradient("Smoke-Impacted Days",
                      low = "#3964C3", high = "#dc2430") +
  # add county path on top
  geom_path() +
  facet_wrap(~smk_cut_method) +
   theme(strip.background = element_blank(),
       strip.text.x = element_blank()) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Oregon: Counties Affected by Smoke PM2.5 > 10µg/m^3") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(colour=NA, fill=NA)) +
  geom_text(data = centroids, aes(long, lat, label = rownames(centroids), group=NULL), size = 2)

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- inhaler_ndc_df %>%
  filter(!is.na(geo_smk_pm_zip)) %>% 
  # only look at outcomes
  filter(outcome == 1) %>%
  # add another row that makes sure there is a person <15 in the dataframe
  # tricking xtabs to produce a 0 cell for the outcome for age <15
  add_row(outcome = 0, age_ind = 0)

@@

### Outcomes of four kinds

```{r category, echo = F, message = F, warning = F, results='asis'}
read_path <- paste0("../data_new/medication/oregon_beta_ndc_index.csv")
inhaler_index <- read_csv(read_path)

# dataframe to loop through
inhaler_index_df <- data.frame(inhaler_index) %>%
  select(personkey, index)

cate <- inhaler_index_df %>%
  filter(personkey %in% inhaler_ndc$personkey) %>%
  unique()

inhaler_ndc_index_df <- inhaler_ndc_df %>%
  left_join(cate, by="personkey")

asthma_df <- inhaler_ndc_index_df %>%
  filter(index=="asthma")

copd_df <- inhaler_ndc_index_df %>%
  filter(index=="copd")

both_df <- inhaler_ndc_index_df %>%
  filter(index=="both")

other_df <- inhaler_ndc_index_df %>%
  filter(index=="other")

df_list <- list(asthma_df, copd_df, both_df, other_df)

outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')

# create an empty list to row bind dataframes together
datalist <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

#str(combined_point_est_df)

tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
           caption = "Number of cases for each outcome observed from May 1st to September 30st, 2013",
           # column headers
           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                      "(%)", "Female", "(%)", "Male", "(%)"),
           # column spanner
           cgroup = c("","Age Category", "Sex"), 
           n.cgroup = c(2, 6, 4),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc" # column alignment,
            ) # end table

@@

The plot shows that four different categories

```{r category plot, warning =F, echo = F, results='asis'}
df_list <- list(asthma_df, copd_df, both_df, other_df)

outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[75])
  # outcome name
  outcome_name <- outcome_list[i]

  covariates_df <- df_to_loop[, c(1:26, 74:84)]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(91, 93)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(geo_smk_pm_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 9, byrow = T)

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

# set row number to fill
row_n <- 1

# only run the model if the dataframe has observations
if(nrow(df_analysis) != 0){
  # conditional logistic regression model
  mod <- clogit(outcome ~ df_analysis[[38]] + wrf_temp_zip + strata(personkey), df_analysis)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)      
  # create else statement that fills matrix with missing so I still have the row
  # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list
          point_estimates[row_n, 3] <- 0
          point_estimates[row_n, 4] <- 0
          point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(1, 4)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
# ggsave("../plot_new/fig_both_er_ur.pdf", plot = print_plot, 
#        width = 12, height = 8, units = "in")

@@

```{r input, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

```{r input category, echo = FALSE, message=FALSE, warning=FALSE}
df_list <- list(asthma_df, copd_df, both_df, other_df)

outcome_list <- c('Asthma', 'COPD', 'Both', 'Other')

# create an empty list to row bind dataframes together
datalist_dml <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

# datalist_dml[[i]] <- data.frame(df_list[i]) %>%
  asthma_dml <- asthma_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)

asthma_dml$day[which(asthma_dml$date=="2013-05-27")] <- "weekend"
asthma_dml$day[which(asthma_dml$date=="2013-07-04")] <- "weekend"
asthma_dml$day[which(asthma_dml$date=="2013-09-02")] <- "weekend"

datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-05-27")] <- "weekend"
datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-07-04")] <- "weekend"
datalist_dml[[i]]$day[which(datalist_dml[[i]]$date=="2013-09-02")] <- "weekend"
}

for(i in 1:4){
  i <- 1

  inhaler_geo_lag <- asthma_dml %>%
  select(159, 160:166) # GWR-method

inhaler_temp_lag <- asthma_dml %>%
  select(93, 150:156) # GWR-method

dml_df <- datalist_dml[[i]]

fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
              asthma_dml)

}

@@

### Find the smallest of AIC with df from 3 to 6 (10)

```{r, message=FALSE, echo=FALSE, warning=FALSE}
